Field and experimental data together with computational

### 2022-03-08 manuscript_PLOS/manuscript_PLOS.Rmd
Add last referece and URL of th repo in the manuscript

---
title: Field and experimental data together with computational 
  models reveal how cleaner fish adjust decisions  in a biological market
author:
  - name: Andrés E. Quiñones 
    email: andresqp@gmail.com
    affiliation: Institute of Biology, University of Neuchatel
    corresponding: andresqp@gmail.com
  - name: Zegni Triki
    email: zegni.triki@gmail.com
    affiliation: Department of Zoology, Stockholm University
  - name: Redouan Bshary
    email: redouan.bshary@unine.ch
    affiliation: Institute of Biology, University of Neuchatel
address:
  - code: Institute of Biology, University of Neuchatel
    address: Institute of Biology, Rue Emile-Argand 11, 2000-CH Neuchatel, Switzerland
  - code: Department of Zoology, Stockholm University
    address: Department of Zoology, Svante Arrheniusväg 18 B,  106 91 Stockholm, Sweden,
abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. We illustrate this point by fitting computational models to experimental data on performance by wild-caught cleaner fish *Labroides dimidiatus* in an ephemeral reward task, as well as cleaner and client fish densities from the locations of capture. Using Bayesian statistics to fit the model parameters to performance data revealed that cleaner fish most likely estimate future consequences of an action, while it appears unlikely that the removal of the ephemeral reward acts as psychological punishment (negative reinforcement). Incorporating future consequences also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement mechanism. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinnings of cognitive performance.

@@

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering the same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish has eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally referred as the ephemeral reward task. 
Individuals are assessed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. 
This between variation in performance, and primates' known cognitive abilities
$\color{red}{\text{ref}}$, shows that performance in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this between species variation suggests
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

@@

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites, we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centred around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equal to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used an 
uninformative prior. Once we decided whether the new parameter set 
would replace the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $1000$ samples of each chain as burn in, 
and after that, we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
$\color{red}{\text{include citation}}$

>>>>

---
title: Field and experimental data together with computational 
  models reveal how cleaner fish adjust decisions  in a biological market
author:
  - name: Andrés E. Quiñones 
    email: andresqp@gmail.com
    affiliation: a
    corresponding: andresqp@gmail.com
  - name: Zegni Triki
    email: zegni.triki@gmail.com
    affiliation: b
  - name: Redouan Bshary
    email: redouan.bshary@unine.ch
    affiliation: Institute of Biology, University of Neuchatel
address:
  - code: a
    address: Institute of Biology, University of Neuchatel, Rue Emile-Argand 11, 2000-CH Neuchatel, Switzerland
  - code: b
    address: Department of Zoology, Stockholm University, Svante Arrheniusväg 18 B,  106 91 Stockholm, Sweden,
abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. We illustrate this point by fitting computational models to experimental data on performance by wild-caught cleaner fish *Labroides dimidiatus* in an ephemeral reward task, as well as cleaner and client fish densities from the locations of capture. Using Bayesian statistics to fit the model parameters to performance data revealed that cleaner fish most likely estimate future consequences of an action, while it appears unlikely that the removal of the ephemeral reward acts as psychological punishment (negative reinforcement). Incorporating future consequences also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement mechanism. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinnings of cognitive performance.

@@

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering the same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish has eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally referred as the ephemeral reward task. 
Individuals are assessed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. 
This between variation in performance, and primates' known cognitive abilities
[@reader_Evolution_2001], shows that performance in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this between species variation suggests
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

@@

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites, we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centred around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equal to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used an 
uninformative prior. Once we decided whether the new parameter set 
would replace the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $1000$ samples of each chain as burn in, 
and after that, we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
<https://github.com/andreseqp/learnDatamodel>

### 2022-03-08 manuscript_PLOS/manuscript_PLOS.tex
Add last referece and URL of th repo in the manuscript

\newcommand{\getAff}[1]{
  \getIndex{#1}{Institute of Biology, University of Neuchatel,Department
of Zoology, Stockholm University}
}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Field and experimental data together with computational
models reveal how cleaner fish adjust decisions in a biological
market} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Andrés E. Quiñones\textsuperscript{\getAff{Institute of Biology,
University of Neuchatel}}\textsuperscript{*},
Zegni Triki\textsuperscript{\getAff{Department of Zoology, Stockholm
University}},
Redouan Bshary\textsuperscript{\getAff{Institute of Biology, University
of Neuchatel}}\\
\bigskip
\textbf{\getAff{Institute of Biology, University of Neuchatel}}Institute
of Biology, Rue Emile-Argand 11, 2000-CH Neuchatel, Switzerland\\
\textbf{\getAff{Department of Zoology, Stockholm University}}Department
of Zoology, Svante Arrheniusväg 18 B, 106 91 Stockholm, Sweden,\\
\bigskip
* Corresponding author: andresqp@gmail.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
While it is generally straightforward to quantify individual performance
in cognitive experiments, identifying the underlying cognitive processes
remains a major challenge. Often, different mechanistic underpinnings
yield similar performances, and Lloyd Morgan's cannon warrants
acceptance of the simpler explanation. Alternatively, when the different
mechanisms interact with environmental conditions, variation in
performance across environments might allow to statistically infer the
mechanism responsible. We illustrate this point by fitting computational
models to experimental data on performance by wild-caught cleaner fish
\emph{Labroides dimidiatus} in an ephemeral reward task, as well as
cleaner and client fish densities from the locations of capture. Using
Bayesian statistics to fit the model parameters to performance data
revealed that cleaner fish most likely estimate future consequences of
an action, while it appears unlikely that the removal of the ephemeral
reward acts as psychological punishment (negative reinforcement).
Incorporating future consequences also yields performances that can be
considered the result of locally optimal decision-rules, in contrast to
the negative reinforcement mechanism. We argue that the combination of
computational models with data is a powerful tool to infer the
mechanistic underpinnings of cognitive performance.

@@

When researchers aimed at testing wild-caught cleaner fish, as well as
individuals from other species, abilities to prioritize visitor over
resident clients in a lab-based paradigm, they used Plexiglas plates of
different colours and shapes offering the same amount of food as
surrogates for visitor and resident clients. One plate is made to behave
like a resident, i.e.~it remains until the cleaner fish has eaten off
it. The other plate was made to behave like a visitor, i.e.~it is
removed if not inspected first. This laboratory paradigm originally
inspired by the cleaning market of cleaner fish is more generally
referred as the ephemeral reward task. Individuals are assessed in their
capacity to develop a preference for the ephemeral option, which gives
them access to two reward units. Cleaners learned to prefer the visitor
plate and hence obtained the double amount of food {[}12{]}.
Furthermore, adult cleaner fish outperformed various primates as well as
rats and pigeons in this original version of the ephemeral reward task
{[}13,14{]}; yet the African grey parrots solve this task as well
{[}15{]}. This between variation in performance, and primates' known
cognitive abilities \(\color{red}{\text{ref}}\), shows that performance
in the task cannot be readily explained by classifying species as having
high and low cognitive abilities. Instead, this between species
variation suggests that solving the ephemeral reward task requires a set
of cognitive mechanisms that are not present in all the tested taxa, or
were not used under the experimental set-up.

@@

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We
started the chain with random values for the parameters of interest,
then ran the computational model once for every reef site using as input
the ecological explanatory variables. The model outputs the probability
of choosing the visitor for each location \(p_i\), where \(i\) is the
index for the 12 locations. Assuming a binomial distribution, the
probability that each of the cleaner fish 20 choices in the ephemeral
reward task was generated by the model is given by
\(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\); where \(k_{ij}\) is
the number of times that cleaner fish \(j\) in reef site \(i\) chose the
visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing
over all the cleaner fish and reef sites, we obtained the log-likelihood
of the data given the model and parameters. We then proposed a new set
of parameters drawn from a uniform distribution centred around the old
parameter set. The amplitude of the uniform distribution used for each
parameter can be found in Table \ref{tab:param}. Subsequently, we ran
the model and calculated the likelihood with the new parameter set. We
then used the ratio of the two likelihoods to choose which parameter set
to keep. New parameter sets with a higher likelihood than the old set
replaced old ones, and those with a lower likelihood replaced current
ones with a probability equal to the log-likelihood ratio. Given that we
only used the likelihoods in the decision, we used an uninformative
prior. Once we decided whether the new parameter set would replace the
old one, we ran the model again to sample the likelihood distribution of
the parameter set. We then started the cycle again by proposing a new
set of parameters and repeated the process for \(1e^5\) steps. We ran 5
independent chains, discarded the first \(1000\) samples of each chain
as burn in, and after that, we kept 1 in every 100 samples to avoid
autocorrelation. The collection of parameters sets kept in all chains
approximates the posterior distribution. We coded the model as well as
the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R {[}32{]}. All codes are accessible at
\(\color{red}{\text{include citation}}\)

>>>>

\newcommand{\getAff}[1]{
  \getIndex{#1}{a,b}
}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Field and experimental data together with computational
models reveal how cleaner fish adjust decisions in a biological
market} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Andrés E. Quiñones\textsuperscript{\getAff{a}}\textsuperscript{*},
Zegni Triki\textsuperscript{\getAff{b}},
Redouan Bshary\textsuperscript{\getAff{Institute of Biology, University
of Neuchatel}}\\
\bigskip
\textbf{\getAff{a}}Institute of Biology,University of Neuchatel ,Rue
Emile-Argand 11, 2000-CH Neuchatel, Switzerland\\
\textbf{\getAff{b}}Department of Zoology, Stockholm University, Svante
Arrheniusväg 18 B, 106 91 Stockholm, Sweden,\\
\bigskip
* Corresponding author: andresqp@gmail.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
While it is generally straightforward to quantify individual performance
in cognitive experiments, identifying the underlying cognitive processes
remains a major challenge. Often, different mechanistic underpinnings
yield similar performances, and Lloyd Morgan's cannon warrants
acceptance of the simpler explanation. Alternatively, when the different
mechanisms interact with environmental conditions, variation in
performance across environments might allow to statistically infer the
mechanism responsible. We illustrate this point by fitting computational
models to experimental data on performance by wild-caught cleaner fish
\emph{Labroides dimidiatus} in an ephemeral reward task, as well as
cleaner and client fish densities from the locations of capture. Using
Bayesian statistics to fit the model parameters to performance data
revealed that cleaner fish most likely estimate future consequences of
an action, while it appears unlikely that the removal of the ephemeral
reward acts as psychological punishment (negative reinforcement).
Incorporating future consequences also yields performances that can be
considered the result of locally optimal decision-rules, in contrast to
the negative reinforcement mechanism. We argue that the combination of
computational models with data is a powerful tool to infer the
mechanistic underpinnings of cognitive performance.

@@

When researchers aimed at testing wild-caught cleaner fish, as well as
individuals from other species, abilities to prioritize visitor over
resident clients in a lab-based paradigm, they used Plexiglas plates of
different colours and shapes offering the same amount of food as
surrogates for visitor and resident clients. One plate is made to behave
like a resident, i.e.~it remains until the cleaner fish has eaten off
it. The other plate was made to behave like a visitor, i.e.~it is
removed if not inspected first. This laboratory paradigm originally
inspired by the cleaning market of cleaner fish is more generally
referred as the ephemeral reward task. Individuals are assessed in their
capacity to develop a preference for the ephemeral option, which gives
them access to two reward units. Cleaners learned to prefer the visitor
plate and hence obtained the double amount of food {[}12{]}.
Furthermore, adult cleaner fish outperformed various primates as well as
rats and pigeons in this original version of the ephemeral reward task
{[}13,14{]}; yet the African grey parrots solve this task as well
{[}15{]}. This between variation in performance, and primates' known
cognitive abilities {[}\textbf{reader\_Evolution\_2001?}{]}, shows that
performance in the task cannot be readily explained by classifying
species as having high and low cognitive abilities. Instead, this
between species variation suggests that solving the ephemeral reward
task requires a set of cognitive mechanisms that are not present in all
the tested taxa, or were not used under the experimental set-up.

@@

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We
started the chain with random values for the parameters of interest,
then ran the computational model once for every reef site using as input
the ecological explanatory variables. The model outputs the probability
of choosing the visitor for each location \(p_i\), where \(i\) is the
index for the 12 locations. Assuming a binomial distribution, the
probability that each of the cleaner fish 20 choices in the ephemeral
reward task was generated by the model is given by
\(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\); where \(k_{ij}\) is
the number of times that cleaner fish \(j\) in reef site \(i\) chose the
visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing
over all the cleaner fish and reef sites, we obtained the log-likelihood
of the data given the model and parameters. We then proposed a new set
of parameters drawn from a uniform distribution centred around the old
parameter set. The amplitude of the uniform distribution used for each
parameter can be found in Table \ref{tab:param}. Subsequently, we ran
the model and calculated the likelihood with the new parameter set. We
then used the ratio of the two likelihoods to choose which parameter set
to keep. New parameter sets with a higher likelihood than the old set
replaced old ones, and those with a lower likelihood replaced current
ones with a probability equal to the log-likelihood ratio. Given that we
only used the likelihoods in the decision, we used an uninformative
prior. Once we decided whether the new parameter set would replace the
old one, we ran the model again to sample the likelihood distribution of
the parameter set. We then started the cycle again by proposing a new
set of parameters and repeated the process for \(1e^5\) steps. We ran 5
independent chains, discarded the first \(1000\) samples of each chain
as burn in, and after that, we kept 1 in every 100 samples to avoid
autocorrelation. The collection of parameters sets kept in all chains
approximates the posterior distribution. We coded the model as well as
the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R {[}32{]}. All codes are accessible at
\url{https://github.com/andreseqp/learnDatamodel}
